An FMI-Based Initialization Plugin for INTO-CPS Maestro 2

### 2020-07-03 background.tex
Updated plots

\begin{definition}[Fixed Point Computation]\label{def:fixedpoint}
The $\fixedpoint{l}$ represents an order sequence of the setting or getting of all input and output variables of a given SCC $l$, see \cref{def:loop} for a definition of SCC.
\end{definition}

@@

>>>>

\begin{definition}[Fixed Point]\label{def:fixedpoint}
The $\fixedpoint{l}$ represents an order sequence of the setting or getting of all input and output variables of a given SCC $l$, see \cref{def:loop} for a definition of SCC.
The $\fixedpoint{l} \subseteq \bigcup_{c \in \fmus} \set{\fget{c},\fset{c}}$ 
\end{definition}

@@

It should be noted that a trivial SCC (\see \cref{def:loop}) is only a single get or set action and is not regarded as a fixed point outside \cref{def:initialization}, but just a simple computation.

### 2020-07-03 casestudy.tex
Updated plots

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\textwidth]{images/init_state_0_sim.png}
    \caption{Simulation results when position and velocity of the chassis mass is zero. \claudio{Simon, please improve plot (I have exported an svg and also the csv with the simulation data).}}
    \label{fig:init_state_0_sim}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.8\textwidth]{images/init_state_correct_sim.pdf}
    \caption{Simulation results starting from a correct initial state (a steady state). \claudio{Simon, please improve plot (I have exported an svg and also the csv with the simulation data). Perhaps even merging the two plots might make sense.}}
    \label{fig:init_state_correct_sim}
\end{figure}

>>>>

\begin{figure}[htb]
\centering
\begin{tikzpicture}
    \begin{axis}[
        at={(0,0)},
        width= 12cm,
        height = 8cm,
        ymin=-0.2,
        ymax=0.05,
        xmin=0,
        xmax=4.1,
        grid=both,
        grid style={line width=.1pt, draw=gray!10},
        major grid style={line width=.2pt,draw=gray!50},
        axis lines=middle,
        minor tick num =5,
        axis line style={latex-latex},
        ticklabel style={font=\tiny,fill=white},
        minor tick style={draw=none},
        minor grid style={thin,color=black!10},
        ylabel= Car chassis position (m),
        xlabel= Time (s),
        tick align=outside,
        x label style={at={(axis description cs:0.5,-0.1)},anchor=north, color=blue!50!cyan},
        y label style={at={(axis description cs:-0.1,.5)},rotate=90,anchor=south, color=blue!50!cyan},
        x tick label style={
            /pgf/number format/assume math mode, font=\sf\scriptsize}
        ]
    \addplot [line width=2.5pt, red] table [x= a, y=b, col sep=comma] {incorrect.csv};
\end{axis}
\end{tikzpicture}
    \caption{Simulation results when position and velocity of the chassis mass is zero.}
    \label{fig:init_state_0_sim}
\end{figure}

@@

\begin{figure}[htb]
\centering
\begin{tikzpicture}
    \begin{axis}[
        %at={(0,0)},
        width= 12cm,
        height = 8cm,
        ymin=-0.18,
        ymax=-0.05,
        xmin=0,
        xmax=4.1,
        grid=both,
        grid style={line width=.1pt, draw=gray!10},
        major grid style={line width=.2pt,draw=gray!50},
        axis lines=middle,
        minor tick num =5,
        axis line style={latex-latex},
        ticklabel style={font=\tiny,fill=white},
        minor tick style={draw=none},
        minor grid style={thin,color=black!10},
        ylabel= Car chassis position (m),
        xlabel= Time (s),
        tick align=outside,
        x label style={at={(axis description cs:0.5,-0.1)},anchor=north, color=blue!50!cyan},
        y label style={at={(axis description cs:-0.1,.5)},rotate=90,anchor=south, color=blue!50!cyan},
        x tick label style={
            /pgf/number format/assume math mode, font=\sf\scriptsize}
            ]
    \addplot [line width=2.5pt, red] table [x=a, y=b, col sep=comma] {correct.csv};
\end{axis}
\end{tikzpicture}
    \caption{Simulation results starting from a correct initial state (a steady state).}
    \label{fig:init_state_correct_sim}
\end{figure}

### 2020-07-03 initializationorder.tex
Updated plots

\section{Calculation of an Initialization Order}\label{sc:initilization}
The FMI specification defines certain information about the initialization order described through different states of a co-simulation. The initialization phase covers the two states (in chronological order) defined in the specification:
\begin{itemize}
    \item \textit{Instantiated}
    \item \textit{Initialization Mode}
\end{itemize}
In each of the two states, different groups of FMU variables and parameters are potentially assigned a value. The groups are defined by FMI based on rules of the characteristics of the variables. These rules have been extracted as predicates and used in the implementation. 
Some groups consist of variables and parameters whose value does not depend on other variables. These can be set before entering the \textit{Initialization Mode}. Since these variables have no connections to other FMU variables - meaning they are not represented in the graph of \cref{def:initialization_graph}, the order is insignificant. 
The operations are on the variables of each FMU are grouped to perform the fewest possible operations in the initialization.

@@

%\subsubsection{Calculation of an initialization order}
The topological ordering or the strongly connected components of the graph defined in \cref{def:initialization_graph} is the initialization order of the interconnected FMU variables. 
The non-trivial strongly connected components are algebraic loops of the system. The trivial ones are standard interconnected FMU variables, whose port operation should only be performed only once during the initialization procedure.
The calculation of an initialization order is performed in linear time based on the number of both external and internal connections using Tarjan's algorithm \cite{tarjan_1972}. 
%This algorithm is selected due to its properties. It solves two of the central issues of the initialization-phase of the co-simulation.
%\begin{itemize}
%    \item Identifies algebraic loops between interconnected variables (strongly connected components)
%    \item Performs a topological sorting of the Strongly Connected Components
%\end{itemize}
The algorithm is well-established, and there exist formal proofs of its correctness and properties\cite{stefanMerz}. 
% The algorithm is among the most efficient graph algorithms for accomplishing the defined goals.
% Tarjan's algorithm is performing a topological sorting of the strongly connected components of a graph. Moreover, it can handle both graphs with and without algebraic loops.

%\subsubsection{Handling of algebraic loops}
As described earlier sections, it is essential to handle the algebraic loops by a particular initialization strategy since they otherwise would invalidate the result of the co-simulation. The strategy for managing algebraic loops is to identify and initialize them using a fixed point iteration until convergence. Since convergence is not guaranteed, this property is monitored using \cref{def:convergence}. If convergence is not established within a finite number of iteration, the co-simulation is rejected to avoid running an invalid simulation.

>>>>

\section{Calculation of an Initialization Order}\label{sc:initilization}
The FMI specification defines certain information about the initialization order described through different states of a co-simulation. The initialization phase covers the two states (in chronological order) defined in the specification:
\begin{itemize}
    \item \textit{Instantiated}
    \item \textit{Initialization Mode}
\end{itemize}
In each of the two states, different groups of FMU variables and parameters are potentially assigned a value. The groups are defined by FMI based on rules of the characteristics of the variables. These rules have been extracted as predicates and used in the implementation. 
Some groups consist of variables and parameters whose value does not depend on other variables. These can be set before entering the \textit{Initialization Mode}. Since these variables have no connections to other FMU variables - meaning they are not represented in the graph of \cref{def:initialization_graph}, the order is insignificant. 
The setting and getting operations are of each FMU are grouped to perform the fewest possible operations in the initialization.

@@

%\subsubsection{Calculation of an initialization order}
The topological ordering of the strongly connected components of the graph defined in \cref{def:initialization_graph} is the initialization order of the interconnected FMU variables. 
The non-trivial strongly connected components are algebraic loops of the system. The trivial ones are standard interconnected FMU variables, whose port operation should be performed only once during the initialization procedure.
The calculation of an initialization order is performed in linear time based on the number of both external and internal connections using Tarjan's algorithm \cite{tarjan_1972}. 
%This algorithm is selected due to its properties. It solves two of the central issues of the initialization-phase of the co-simulation.
%\begin{itemize}
%    \item Identifies algebraic loops between interconnected variables (strongly connected components)
%    \item Performs a topological sorting of the Strongly Connected Components
%\end{itemize}
The algorithm is well-established, and there exist formal proofs of its correctness and properties\cite{stefanMerz}. 
% The algorithm is among the most efficient graph algorithms for accomplishing the defined goals.
% Tarjan's algorithm is performing a topological sorting of the strongly connected components of a graph. Moreover, it can handle both graphs with and without algebraic loops.

%\subsubsection{Handling of algebraic loops}
As described in earlier sections, it is essential to handle the algebraic loops by a particular initialization strategy since they otherwise would invalidate the result of the co-simulation. The strategy for managing algebraic loops is to identify and initialize them using a fixed point iteration until convergence. Since convergence is not guaranteed, this property is monitored using \cref{def:convergence}. If convergence is not established within a finite number of iteration, the co-simulation is rejected to avoid running an invalid simulation.

### 2020-07-03 realization.tex
Started looking at Claudios comments
Updated plots

\section{Realization of a Maestro 2 Plugin}\label{sc:implementation}
This approach has been realized as a Maestro 2 plugin that generates the \textit{Initialization}-phase of a co-simulation specification expressed in MaBL. It calculates the specification based on a list of FMU-components of a co-simulation scenario, and a specific plugin-configuration to let the user supply initial values of FMU variables with $causality=parameter$. A correction initialization specification can be calculated by the plugin if the co-simulation scenario is adhering to the behavior dictated by definition given in \cref{def:convergence}.

@@

The Prolog code is capable of generating a valid initialization order, but this functionality is not used for several reasons. The main reasons are based on performance considerations and the desire to handle the initialization of co-simulation scenarios containing strongly connected components.

>>>>

\section{Realization of a Maestro 2 Plugin}\label{sc:implementation}
This approach has been realized as a Maestro 2 plugin that generates the \textit{Initialization}-phase of a co-simulation specification expressed in MaBL. It calculates the specification based on a list of FMU-components of a co-simulation scenario, and a specific plugin-configuration to let the user supply initial values of FMU variables with $causality=parameter$. A correction initialization specification can be calculated by the plugin if the co-simulation scenario is adhering to the behavior dictated by the definition given in \cref{def:convergence}.

@@

### 2020-07-03 relatedwork.tex
Updated background section
Updated plots

The work by \cite{EvoraGomez2019a} in of a similar nature to ours. They use Tarjan’s SCC algorithm to generate a sorted DAG of strongly connected components to solve the initialization problem.
Even though their work is very similar to ours, we extend their approach with a verification against the simulation semantics resulting in a formal more sounds approach. However further work will look into further improvements of the current approach.

>>>>

The work by Gomez et al. .\cite{EvoraGomez2019a} is similar to ours. They use Tarjan's SCC algorithm to generate a sorted DAG of strongly connected components to solve the initialization problem.
Even though their work is very similar to ours, we extend their approach with verification against the simulation semantics resulting in a formally more sound approach. However, further work will look into further improvements and formal verification of the current approach.
